---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# isoseabird

<!-- badges: start -->
<!-- badges: end -->

A package to analyse isotopic niche overlaps in seabirds

It provides data and uses functions from other packages to create a workflow. Please see references to refer to the packages.

# Intro

This package contains:

- Data from stable isotopes from red-footed and masked boobies from Isla Clarion

- Eleven functions to organize your data and to calculate Mean distance to centroid and Euclidian distance, adapted from Turner et al. 2010

- Four functions to organize, convert objects and calculate proportion of overlap using the package SIBER

- Six functions to plot ellipses using ggplot and to adapt the theme

## Install package

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("MiriamLL/isoseabird")
```

Load package

```{r}
library('isoseabird')
```

# Data

Contains C and N isotopic blood data from red-footed and masked boobies breeding at Clarion Island

```{r, eval=FALSE}
head(Sula_data)
```

```{r}
Sula_data<-Sula_data
```


# 1. Turner

Contains functions from Turner et al 2010   
A general hypothesis-testing framework for stable isotope ratios in ecological studies  
Ecology 91(8):2227-33

Please refer to [Turner et al. 2010](https://esapubs.org/archive/ecol/E091/157/)  
For details on the script go to: [R Files](https://esapubs.org/archive/ecol/E091/157/suppl-1.htm)

## 1.1. organize_for_Turner

```{r}
SIA_Turner<-organize_for_Turner(SIA_samples=Sula_data,
                                column_C='Carbon',
                                column_N='Nitrogeno',
                                column_group='Species')
```

## 1.2. plot_centroids

```{r, fig.height=5, fig.width=8, fig.align='center'}
plot_centroids(SIA_samples=Sula_data,
               column_C='Carbon',
               column_N='Nitrogeno',
               column_group='Species',
               group1="SulaSula",
               group2="Dactylatra")
```

## 1.3. create_matrix

```{r}
SIA_matrix<-create_matrix(SIA_Turner = SIA_Turner)
```

## 1.4. Calculate_residuals

```{r}
SIA_residuals<-calculate_residuals(SIA_matrix=SIA_matrix,
                                      SIA_group = SIA_group)
```

## 1.5. dispersion_measures

```{r}
SIA_dispersion<-calculate_dispersion(SIA_residuals = SIA_residuals,
                                     SIA_group=SIA_group)
```

## 1.6. dispersion statistics

```{r}
SIA_tabledispersion<-dispersion_statistics(SIA_dispersion=SIA_dispersion,
                                           SIA_group=SIA_group)
SIA_tabledispersion
```



## 1.7. calculate_means

```{r}
SIA_means<-calculate_means(SIA_matrix,SIA_group)
SIA_means
```

```{r}
SIA_meandif<-as.vector(dist(as.matrix(SIA_means)))
SIA_meandif
```

## 1.8. contrasting differences


```{r}
SIA_dispersiondif<-contrast_differences(SIA_tabledispersion = SIA_tabledispersion)
SIA_dispersiondif
```

```{r}
SIA_dispersiondif$mdc
```

## 1.9. permutation procedure

```{r, eval=FALSE}
SIA_permutation<-run_permutation(SIA_matrix=SIA_matrix, 
                                 SIA_group=SIA_group,
                                 SIA_dispersiondif = SIA_dispersiondif,
                                 permutation_runs=999)
```


## 1.10. calculate_pvalues

```{r}
SIA_pvalue<-calculate_pvalues(SIA_permutation=SIA_permutation)
SIA_pvalue
```


## 1.11. hotellings

*Hotellings revealed that Euclidian distance between centroids differed significantly from zero for contrast between guilds (Turner et al. 2010)*

```{r}
SIA_hotellings<-calculate_hotellings(SIA_means=SIA_means,
                               SIA_group=SIA_group,
                               SIA_matrix=SIA_matrix)
SIA_hotellings
```



# 2. SIBER

Creates work flow to use the package SIBER to calculate ellipses overlaps.

Refer to the paper: [Jackson et al. 2011](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/j.1365-2656.2011.01806.x)  
Check for updates on: [SIBER - last update 2023](https://cran.r-project.org/web/packages/SIBER/vignettes/Introduction-to-SIBER.html)

## 2.1. organize_for_SIBER

```{r}
SIBER_SIA<-organize_for_SIBER(SIA_samples=Sula_data,
                              column_C='Carbon',
                              column_N='Nitrogeno',
                              column_group='Species')
```

## 2.2. convert_object

```{r}
SIBER_object<-convert_object(SIBER_SIA=SIBER_SIA)
str(SIBER_object)
```

```{r}
SIBER_group<-SIBER::groupMetricsML(SIBER_object)
SIBER_group
```


## 2.3. overlap_ellipses

This function organizes the data and uses the function [maxLikOverlap](https://search.r-project.org/CRAN/refmans/SIBER/html/maxLikOverlap.html)

```{r}
SIBER_overlap_50<-areas_overlap(SIBER_object = SIBER_object,
                              group1<-'SulaSula',
                              group2<-'Dactylatra',
                              overlap = 0.50)
SIBER_overlap_50
```

```{r}
SIBER_overlap_95<-areas_overlap(SIBER_object = SIBER_object,
                              group1<-'SulaSula',
                              group2<-'Dactylatra',
                              overlap = 0.95)
SIBER_overlap_95
```

## 2.4. prop_overlap

```{r}
prop_overlap_50<-proportion_overlap(area_overlap = SIBER_overlap_50)
prop_overlap_50
```

```{r}
prop_overlap_95<-proportion_overlap(area_overlap = SIBER_overlap_95)
prop_overlap_95
```


# 3. Figure


```{r}
Sula_data<-Sula_data
```

## 3.1. ggplot_ellipses

```{r, fig.height=2}
SIA_plot<-ggplot_ellipse(SIA_samples<-Sula_data,column_C<-'Carbon',column_N<-'Nitrogeno',
                         column_group<-'Species')
SIA_plot
```


## 3.2. adapt_axis

```{r, fig.height=2}
SIA_plot_axis<-adapt_axis(SIA_plot=SIA_plot,
                            C_breaks=c(-20,-19,-18,-17,-16,-15,-14),
                            C_limits=c(-20,-14),
                            N_breaks=c(14,15,16,17,18),
                            N_limits=c(13.5,18.1))
SIA_plot_axis
```


## 3.3. adapt_theme

```{r, fig.height=2}
SIA_plot_theme<-adapt_theme(SIA_plot=SIA_plot_axis)
SIA_plot_theme
```

## 3.4. adapt_legend

```{r, fig.height=2}
SIA_plot_legend<-adapt_legend(SIA_plot=SIA_plot_theme,
                       xlabel_adjust=-19.9,
                       ylabel_adjust=17.9,
                       label1='Masked booby',
                       label2='Red-footed booby',
                       color1='#233d4d',
                       color2='#2a9d8f')
SIA_plot_legend
```

## 3.5. adapt_colors

```{r, fig.height=5}
SIA_plot_colors<-adapt_colors(SIA_plot=SIA_plot_legend,
                              color1='#233d4d',
                              color2='#2a9d8f',
                              column_group='Species')
SIA_plot_colors
```


# 4. Overview

## Mean distance to centroid

```{r}
SIA_dispersiondif$mdc
```

```{r}
SIA_pvalue
```


## Euclidian distance

```{r}
SIA_hotellings
```

## Proportion of overlap

```{r}
prop_overlap_50
prop_overlap_95
```

## Figure

```{r}
SIA_plot_colors
```

## Export

```{r, eval=FALSE}
ResultsFolder<-here::here("3Results")
SIA_plot_colors
ggsave("Fig2.jpg",
       path = ResultsFolder,
       width = 10, height = 10, 
       units = "in") 
```


# 5. References


[Hammerschlag-Peyer et al. 2011](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0027104)

[Turner et al. 2010](https://esapubs.org/archive/ecol/E091/157/)  
[R Files](https://esapubs.org/archive/ecol/E091/157/suppl-1.htm)

[Jackson et al. 2011](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/j.1365-2656.2011.01806.x)  
[SIBER - last update 2023](https://cran.r-project.org/web/packages/SIBER/vignettes/Introduction-to-SIBER.html)  
[SIA in ggplot2](https://cran.r-project.org/web/packages/SIBER/vignettes/Plot-SIA-ggplot2.html)
